---
output: pdf_document
---
#Exploring the relationship between a set of variables and miles per gallon (MPG) 

##Executive Summary  
In this analysis I used the motor trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, I am interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome) and particularly interested in the following two questions,   
                        So,I do linear regression on all variables and exploring relationship between miles per gallon with other variables. I choose those variables that have impact on our model and left others models. Then, I confirm my result by various measures, like regression plot and hypothesis testing. I came up with result that there is quite difference between automatic or manual transmission better for MPG. 

##Data Processing
Firstly I grab libraries and dataset needed for that analysis.

```{r}
library(ggplot2)
data(mtcars)
data<-mtcars
```

##Exploratory Analysis
After that I do little bit of exploratory analysis for understanding, visually seeing and exploring relationships. And see on appendix for its figure 1.1

```{r}
str(mtcars)
colnames(mtcars)
```

##Model design and selection
Firstly I do multivariate regression to explore relationship between all variables and then choosing those that have quite impact on mpg and using backward-elimination strategy to eliminate the unrelated variables one-at-a-time.
```{r}
temp<-summary(lm(data=data,mpg~.-1))$coeff
temp[order(temp[,"Pr(>|t|)"]),][1:3,]
```
and then after selection top three that have impact and choosing wt, qsec, am and neglection others. Then fitting models and seeing quite improvment in p-value

```{r}
ImprovedData<-mtcars[,c("mpg","am","qsec","wt")]
temp<-summary(lm(data=ImprovedData,mpg~.-1))$coeff
temp[order(temp[,"Pr(>|t|)"]),]
```

and coefficients for my model will be 

```{r}
fit<-lm(data=ImprovedData,mpg~.-1)
fit
```

and then quantifying the difference by the following graph,
```{r}
par(mfrow = c(1, 1)); x<-mtcars$mpg; y<-mtcars$am+mtcars$wt+mtcars$qsec
plot(x,y,frame = FALSE,col=mtcars$am+3,cex=1,bg = "salmon")
abline(h = mean(y[mtcars$am==0]), lwd = 3,col="green"); abline(h = mean(y[mtcars$am==1]), lwd = 3,col="blue")
fit1<-lm(I(y[mtcars$am==0])~I(x[mtcars$am==0])); abline(fit1,col="green")
fit2<-lm(I(y[mtcars$am==1])~I(x[mtcars$am==1])); abline(fit2,col="blue")
abline(lm(y~x),lwd=2)
legend("topright", inset=.05, title="Transmisison Type",c("Automatic","Manual"), fill=c("Blue","Green"), horiz=TRUE)
```

for seeing relation among them as, see in appendix for figure 1.2,and then checking validation of my model,
For verifying our model, see various error measures in appendix figure 1.3

##Conclusion

So by looking at my model, it is quite vivid that there is quite difference between automatic or manual transmission better for MPG. And doing t test on it for further verification.

```{r}
coeff <-summary(lm(mpg~.-1, data = ImprovedData))$coeff
CInterval <- coeff["am", 1] + c(-1, 1) * qt(0.975, df = lm(mpg~.-1, data = ImprovedData)$df) # coeff["am",2]
CInterval
```
So, looing at the 95% of our model, it suggests that changing from automatic to manual transmission give a 2.2 to 6.39 increase in miles per gallon for the cars. So, in a nutshell, manual transmission performance is quite good as compagreen to automatic transmission.

##Appendix
Figure 1.1 for exploratory analysis is below
```{r}
p <- ggplot(mtcars, aes(factor(am), mpg))
p + geom_boxplot() + geom_jitter()+ggtitle("MPG VS Transmission type")+xlab("Transmission type")
```
Figure 1.2  seeing relation among varialbles is below

```{r}
pairs(ImprovedData, panel = panel.smooth, main = "comparisons",col=3+ImprovedData$am)
```

Figure 1.3 for various error measures is below

```{r}
par(mfrow = c(2, 2)); par(mar=c(3.1,2.1,1.1,1.1))
fit <- lm(mpg ~ . -1, data = ImprovedData); plot(fit)
```
